Supplementary Material for: A Rough Guide to Pre-processing High-Frequency Animal Tracking Data

### 2021-02-11 _02_filtering_data.Rmd
new section
style code

We prepare the data for plotting.

Histogram of speed

```{r}
fig_hist_speed =
  ggplot()+
  geom_histogram(
    data = data_copy,
    aes(x = in_speed,
        fill = ifelse(in_speed > 0.05, "outlier", "good")),
    position = position_identity(),
    bins = 60
  )+
  geom_histogram(
    data = data_copy[500:800, ],
    aes(x = in_speed),
    fill = pal[4],
    bins = 60
  )+
  geom_vline(
    xintercept = 0.05,
    # size = 0.2,
    lty = 2
  )+
  annotate(
    geom = "text",
    x = 0.11,
    y = 150,
    label = "Threshold",
    size = 4,
    angle = 90
  )+
  scale_fill_manual(
    values = c(
      "outlier" = pal[1],
      "good" = "grey"
    )
  )+
  scale_y_sqrt()+
  scale_x_log10(
  )+
  theme_test(base_family = "sans")+
  theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(
      face = "bold"
    ),
    plot.background = element_blank()
  )+
  labs(
    title = "(c)",
    x = "Speed"
  )
```

```{r echo=FALSE}
# data plot
fig_outlier_remove <-
  ggplot() +
  geom_path(
    data = data_copy,
    aes(x, y,
        col = ifelse(in_speed >= 0.05 | out_speed >= 0.05, 
                     "outlier_path", "good_path"),
        group = NA),
    lwd = 0.2
  ) +
  geom_point(
    data = data_copy[500:800, ],
    aes(x, y),
    # size = 0.5,
    alpha = 0.6,
    # stroke = 1,
    shape = 4, col = pal[4]
  ) +
  geom_point(
    data = data_copy[!data_copy[500:800, ],
      on = c("x", "y")
    ],
    aes(x, y,
      col = ifelse(in_speed >= 0.03 & out_speed >= 0.03, "outlier", "good"),
      shape = (in_speed >= 0.03 & out_speed >= 0.03),
      size = (in_speed >= 0.03 & out_speed >= 0.03)
    ),
    show.legend = F,
    alpha = 1
  ) +
  geom_path(
    data = data,
    aes(x, y),
    col = "grey20",
    alpha = 1
  ) +
  annotation_custom(
    grob = ggplotGrob(
      fig_hist_speed
    ),
    xmin = 0.80,
    xmax = 1.08,
    ymin = 0.25,
    ymax = 0.6
  )+
  scale_color_manual(values = c(
    "outlier_path" = "red",
    "good_path" = "grey",
    "good" = pal[3], 
    "outlier" = "red")) +
  scale_shape_manual(values = c(1, 4)) +
  scale_size_manual(
    values = c(0.2, 2)
  ) +
  ggthemes::theme_few(
    base_family = "Arial"
  ) +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    legend.position = "none"
  ) +
  coord_equal() +
  theme(plot.background = element_rect(fill = NA))
```

>>>>

### Histogram of speed

```{r}
fig_hist_speed <-
  ggplot() +
  geom_histogram(
    data = data_copy,
    aes(
      x = in_speed,
      fill = ifelse(in_speed > 0.05, "outlier", "good")
    ),
    position = position_identity(),
    bins = 60
  ) +
  geom_histogram(
    data = data_copy[500:800, ],
    aes(x = in_speed),
    fill = pal[4],
    bins = 60
  ) +
  geom_vline(
    xintercept = 0.05,
    # size = 0.2,
    lty = 2
  ) +
  annotate(
    geom = "text",
    x = 0.11,
    y = 150,
    label = "Threshold",
    size = 4,
    angle = 90
  ) +
  scale_fill_manual(
    values = c(
      "outlier" = pal[1],
      "good" = "grey"
    )
  ) +
  scale_y_sqrt() +
  scale_x_log10() +
  theme_test(base_family = "sans") +
  theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(
      face = "bold"
    ),
    plot.background = element_blank()
  ) +
  labs(
    title = "(c)",
    x = "Speed"
  )
```

## Figure outlier removal

```{r echo=FALSE}
# data plot
fig_outlier_remove <-
  ggplot() +
  geom_path(
    data = data_copy,
    aes(x, y,
      col = ifelse(in_speed >= 0.05 | out_speed >= 0.05,
        "outlier_path", "good_path"
      ),
      group = NA
    ),
    lwd = 0.2
  ) +
  geom_point(
    data = data_copy[500:800, ],
    aes(x, y),
    # size = 0.5,
    alpha = 0.6,
    # stroke = 1,
    shape = 4, col = pal[4]
  ) +
  geom_point(
    data = data_copy[!data_copy[500:800, ],
      on = c("x", "y")
    ],
    aes(x, y,
      col = ifelse(in_speed >= 0.03 & out_speed >= 0.03, "outlier", "good"),
      shape = (in_speed >= 0.03 & out_speed >= 0.03),
      size = (in_speed >= 0.03 & out_speed >= 0.03)
    ),
    show.legend = F,
    alpha = 1
  ) +
  geom_path(
    data = data,
    aes(x, y),
    col = "grey20",
    alpha = 1
  ) +
  annotation_custom(
    grob = ggplotGrob(
      fig_hist_speed
    ),
    xmin = 0.80,
    xmax = 1.08,
    ymin = 0.25,
    ymax = 0.6
  ) +
  scale_color_manual(values = c(
    "outlier_path" = "red",
    "good_path" = "grey",
    "good" = pal[3],
    "outlier" = "red"
  )) +
  scale_shape_manual(values = c(1, 4)) +
  scale_size_manual(
    values = c(0.2, 2)
  ) +
  ggthemes::theme_few(
    base_family = "Arial"
  ) +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    legend.position = "none"
  ) +
  coord_equal() +
  theme(plot.background = element_rect(fill = NA))
```

### 2021-02-11 _03_smoothing_tracks.Rmd
style code

```{r read_sim_data_2}
# read in the data and set the window size variable
data <- fread("data/data_sim.csv")[5000:10000, ]
data[, window_size := NA]
data$speed_in = atl_get_speed(data)

@@

```{r}
# bind list after offset
data_plot <- mapply(function(df, offset) {
  df = copy(df)
  df[, x := x + offset]
  df$speed_in = atl_get_speed(df)
  return(df)
}, list_of_smooths, seq(0.4, 1.25, length.out = 4),
SIMPLIFY = F
)

@@

# add offset to data errors
data_errors = lapply(c(0, seq(0.4, 1.25, length.out = 4)),
                     function(offset) {
  df = copy(data_errors)
  df[,x := x + offset]
  df[, offset := offset]
})

# bind list
data_errors = rbindlist(data_errors)
```

@@

```{r}
pal2 = RColorBrewer::brewer.pal(4, "RdPu")[c(2, 4)]

fig_hist_smooth =
  ggplot()+
  geom_histogram(
    data = data_errors[offset == 0,],
    aes(x = in_speed,
        fill = "errors"),
    bins = 80,
    alpha = 1
  )+
  geom_histogram(
    data = data_plot[window_size %in% c(5, 21)],
    aes(x = speed_in,
        fill = as.factor(window_size),
        group = window_size),
    bins= 80
  )+
  geom_histogram(
    data = data,
    aes(x = speed_in,
        fill = "real"),
    # col = pal[1],
    alpha = 0.8,
    bins = 80
  )+
  scale_x_log10()+
  scale_fill_manual(
    values = c(
      "real" = alpha("grey20", .8),
      "errors" = "grey",
      "5" = pal2[1],
      "21" = pal2[2]
    ),
    labels = c(
      "real" = "real track",
      "errors" = "track w/\nerrors",
      "5" = "K = 5",
      "21" = "K = 21"
    )
  )+
  coord_cartesian(xlim = c(1e-5, .05))+
  theme_test(base_family = "sans")+
  theme(
    legend.position = c(0.2, 0.65),
    axis.text = element_blank(),
    # axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(
      face = "bold"
    ),
    plot.background = element_blank()
  )+
  labs(x= "Speed [log10 scale]",
       fill = "Smoothing",
       title = "(f)")
```

@@

```{r echo=FALSE}
# prepare data to plot
# make list of data to plot
figure_median_smooth <-
  ggplot() +
  geom_path(
    data = data_errors,
    aes(x, y,
        group = offset),
    col = "grey60",
    size = 0.1
  ) +
  geom_point(
    data = data_errors[offset == 0,],
    aes(x, y),
    col = pal[3],
    size = 0.2
  ) +
  geom_path(
    data = data,
    aes(x, y),
    col = "grey20",
    lwd = 0.5
  ) +
  geom_path(
    data = data_plot,
    aes(x, y,
      group = window_size,
      col = window_size
    ),
    show.legend = F,
    lwd = 0.3
  ) +
  coord_equal(
    expand = F,
    ylim = c(0.6, 0.85),
    xlim = c(NA, 2.3),
    ratio = 1.75
  ) +
  annotate(
    geom = "text",
    x = c(0.75, seq(1.1, 2, length.out = 4)),
    y = 0.82,
    label = sprintf("(%s)", letters[seq(5)]),
    fontface = "bold"
  ) +
  scale_colour_distiller(
    palette = "RdPu", direction = 1,
    values = c(-1, 1)
  ) +
  ggthemes::theme_few() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank()
  )

>>>>

```{r read_sim_data_2}
# read in the data and set the window size variable
data <- fread("data/data_sim.csv")[5000:10000, ]
data[, window_size := NA]
data$speed_in <- atl_get_speed(data)

@@

```{r}
# bind list after offset
data_plot <- mapply(function(df, offset) {
  df <- copy(df)
  df[, x := x + offset]
  df$speed_in <- atl_get_speed(df)
  return(df)
}, list_of_smooths, seq(0.4, 1.25, length.out = 4),
SIMPLIFY = F
)

@@

# add offset to data errors
data_errors <- lapply(
  c(0, seq(0.4, 1.25, length.out = 4)),
  function(offset) {
    df <- copy(data_errors)
    df[, x := x + offset]
    df[, offset := offset]
  }
)

# bind list
data_errors <- rbindlist(data_errors)
```

@@

```{r}
pal2 <- RColorBrewer::brewer.pal(4, "RdPu")[c(2, 4)]

fig_hist_smooth <-
  ggplot() +
  geom_histogram(
    data = data_errors[offset == 0, ],
    aes(
      x = in_speed,
      fill = "errors"
    ),
    bins = 80,
    alpha = 1
  ) +
  geom_histogram(
    data = data_plot[window_size %in% c(5, 21)],
    aes(
      x = speed_in,
      fill = as.factor(window_size),
      group = window_size
    ),
    bins = 80
  ) +
  geom_histogram(
    data = data,
    aes(
      x = speed_in,
      fill = "real"
    ),
    # col = pal[1],
    alpha = 0.8,
    bins = 80
  ) +
  scale_x_log10() +
  scale_fill_manual(
    values = c(
      "real" = alpha("grey20", .8),
      "errors" = "grey",
      "5" = pal2[1],
      "21" = pal2[2]
    ),
    labels = c(
      "real" = "real track",
      "errors" = "track w/\nerrors",
      "5" = "K = 5",
      "21" = "K = 21"
    )
  ) +
  coord_cartesian(xlim = c(1e-5, .05)) +
  theme_test(base_family = "sans") +
  theme(
    legend.position = c(0.2, 0.65),
    axis.text = element_blank(),
    # axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(
      face = "bold"
    ),
    plot.background = element_blank()
  ) +
  labs(
    x = "Speed [log10 scale]",
    fill = "Smoothing",
    title = "(f)"
  )
```

@@

```{r echo=FALSE}
# prepare data to plot
# make list of data to plot
figure_median_smooth <-
  ggplot() +
  geom_path(
    data = data_errors,
    aes(x, y,
      group = offset
    ),
    col = "grey60",
    size = 0.1
  ) +
  geom_point(
    data = data_errors[offset == 0, ],
    aes(x, y),
    col = pal[3],
    size = 0.2
  ) +
  geom_path(
    data = data,
    aes(x, y),
    col = "grey20",
    lwd = 0.5
  ) +
  geom_path(
    data = data_plot,
    aes(x, y,
      group = window_size,
      col = window_size
    ),
    show.legend = F,
    lwd = 0.3
  ) +
  coord_equal(
    expand = F,
    ylim = c(0.6, 0.85),
    xlim = c(NA, 2.3),
    ratio = 1.75
  ) +
  annotate(
    geom = "text",
    x = c(0.75, seq(1.1, 2, length.out = 4)),
    y = 0.82,
    label = sprintf("(%s)", letters[seq(5)]),
    fontface = "bold"
  ) +
  scale_colour_distiller(
    palette = "RdPu", direction = 1,
    values = c(-1, 1)
  ) +
  ggthemes::theme_few() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank()
  )
